---
title: "Appendix 5: Study Diagrams"
author: "Vignette Author"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

A diagram is often very useful for showing where a study took place, how an experiment was designed, how treatments were allocated, or how there are **dependencies** in the data.  Zuur & Ieno (2016) emphasize this ("Step 2: Visualize the experimental design" and "Step 4: Identify the dependency structure in the data"; see Figure 4).

What follows are references to different types of diagrams that you could include in your study.


Below is also an example using a graph of the hierarchical phylogenetic relationships.  

### Site maps: Where a study took place

These papers have examples of site maps for ecological studies; both maps are in the supplemental information.  The focus in more on providing contextual geographic information than the details of the study.

Latta, Brouwer et al.  Avian community characteristics and demographics reveal how conservation value of regenerating tropical dry forest changes with forest age.  PeerJ
https://peerj.com/articles/5217/#supp-1

Latta, Brouwer et al.  Long-term monitoring reveals an avian species credit in secondary forest patches of Costa Rica.  PeerJ
https://peerj.com/articles/3539/#supplemental-information


### Experimental layout maps

Ecological studies sometimes include maps to show the geometric arrangement of study plots or experimental treatments.

Historic disturbance regimes promote tree diversity only under low browsing regimes in eastern deciduous forest
T Nuttle, AA Royo, MB Adams… - Ecological Monographs, 2013 - Wiley Online Library

Long-term biological legacies of herbivore density in a landscape‐scale experiment: forest understoreys reflect past deer density treatments for at least 20 years
T Nuttle, TE Ristau, AA Royo - Journal of Ecology, 2014

Accounting for multilevel data structures in fisheries data using mixed models
T Wagner, DB Hayes, MT Bremigan - Fisheries, 2006


### Dependency structure diagrams

For complex experiments it can be useful to create a diagram that illustrate the hiearchical structure of the data, such as students in class rooms in schools in school districts in counties in states.

#### MCDB / Bio-medicine

Improving basic and translational science by accounting for litter-to-litter variation in animal models SE Lazic, L Essioux - BMC neuroscience, 2013

Four simple ways to increase power without increasing the sample size
Stanley E Lazic Laboratory Animals
https://journals.sagepub.com/doi/full/10.1177/0023677218767478

Analyzing clustered data: why and how to account for multiple observations nested within a study participant? EL Moen, CJ Fricano-Kugler, BW Luikart, AJ O'Malley - Plos one, 2016

#### Ecology

Nested by design: model fitting and interpretation in a mixed model era
H Schielzeth, S Nakagawa - Methods in Ecology and Evolution, 2013

Meta-evaluation of meta-analysis: ten appraisal questions for biologists
S Nakagawa, DWA Noble, AM Senior… - BMC …, 2017 
https://bmcbiol.biomedcentral.com/articles/10.1186/s12915-017-0357-7

Nonindependence and sensitivity analyses in ecological and evolutionary meta-analyses …, M Lagisz, RE O'dea, S Nakagawa - Molecular …, 2017

Zuur & Ieno.  2016.  A protocol for conducting and presenting results of regression‐type analyses
https://besjournals.onlinelibrary.wiley.com/doi/full/10.1111/2041-210X.12577




### Flow charts


#### Clinical trial

Clinical trials often have a flow chart showing how an initial patient population was screen and then allocated to the experimental groups.

https://openi.nlm.nih.gov/detailedresult.php?img=PMC2945952_1471-2474-11-205-2&req=4

#### Meta-analyses

Meta-analyses often have flow charts showing how an initial pool of studies was screen for inclusion in the analysis.

https://openi.nlm.nih.gov/detailedresult.php?img=PMC2559819_1471-2261-8-24-1&req=4


### Coordinated experiments across labs

Resolving discrepant findings on ANGPTL8 in beta-cell proliferation: a collaborative approach to resolving the betatrophin controversy
AR Cox, O Barrandon, EP Cai, JS Rios, J Chavez… - PloS one, 2016


### Dependency structure due to phylogeny

The dependency structure in the mammals milk data is due to phylogeny.  The original paper presents a phylogenetic tree as a figure.  Similarly,  code below is a first attempt to visualize this based on nominal taxonomic categories (class, order, family, genus, species).  Its not very clean yet.  


### Conceptual issues of clustering

The following papers discuss conceptual issues of grouping, clustering etc, but don't have pictures of it.

A solution to dependency: using multilevel analysis to accommodate nested data
E Aarts, M Verhage, JV Veenvliet, CV Dolan… - … neuroscience, 2014

What exactly is 'N'in cell culture and animal experiments?
SE Lazic, CJ Clarke-Williams, MR Munafò - PLoS biology, 2018

The problem of pseudoreplication in neuroscientific studies: is it affecting your analysis? SE Lazic - BMC neuroscience, 2010

Using biological insight and pragmatism when thinking about pseudoreplication
N Colegrave, GD Ruxton - Trends in ecology & evolution, 2017


## Visualizing the structure of the mammals milk data

### Preliminaries

#### Load packages

```{r}
library(data.tree)
library(treemap)
library(igraph)
library(here)
library(stringr)

```

#### Load data

#### Load data in R

```{r}
file. <- "Appendix-2-Analysis-Data_mammalsmilkRA.csv"
path. <- here("/inst/extdata",file.)

milk <- read.csv(path., skip = 3)


milk$ord <- gsub("11","",milk$ord)
```


#### Cleaning

```{r}
genus_spp_subspp_mat <- milk$spp %>% 
         str_split_fixed(" ", n = 3)

milk$genus <- genus_spp_subspp_mat[,1]

```

#### Set up taxonomy 

##### Class, subclass
```{r}
## class - all mammals
milk$class <- "Mammalia"

## Subclass

### everyone but Monotrmes are Theriiformes
#### = Yinotheria
milk$subclass <- "Theriiformes"
milk$subclass[which(milk$ord == "Monotremata")] <- "Yinotheria"
```


##### Cohort 
```{r}
milk$cohort <- "Placentalia"


Marsupialia <- c("Diprotodontia"  #marsupials
                      ,"Didelphimorphia" #opossums
                      ,"Dasyuromorphia"#tasmanian devil)
                      ,"Peramelemorphia" #bandicoots
                  )

milk$cohort[which(milk$ord %in% Marsupialia)] <- "Marsupialia"


milk$cohort[which(milk$ord == "Monotremata")] <- "Australosphenida"


```

##### magnorder
```{r}
milk$magnorder <- "Xenarthra-or-Epitheria"

milk$magnorder[which(milk$ord == "Monotremata")] <- "Monotremata-magnorder"

 Australidelphia <- c("Diprotodontia"  #marsupials
                      #,"Didelphimorphia" #opossums
                      ,"Dasyuromorphia"#tasmanian devil)
                      ,"Peramelemorphia" #bandicoots
                  )



milk$magnorder[which(milk$ord %in% Australidelphia)] <- "Australidelphia"

milk$magnorder[which(milk$ord == "Didelphimorphia")] <- "Ameridelphia"
```


### Use tables to look at dependencies

```{r}
with(milk,
     table(class, subclass))
with(milk,
     table(subclass, cohort))
with(milk,
     table(cohort, magnorder))
with(milk,
     table(ord, magnorder))
```


### Make hierachical graph

### Set up path string

```{r}
#milk$i <- 1:nrow(milk)
milk$pathString <- with(milk,
                        paste(class,  #Mammalia
                              subclass,
                              cohort,
                              magnorder,
                              ord,
                              #fam,
                              #genus,
                              #spp,
                              #i,
                              sep = "/"))
```


#### Set up nodes

```{r}
milk_tree <- as.Node(milk)
```

#### Print hiearchy as text
```{r}
print(milk_tree, limit = 20)
```

#### Plot

Plot branch diagram
```{r}
plot(milk_tree)
```


Plot dendrogram
```{r}
plot(as.dendrogram(milk_tree),
     horiz = T,
     center = T, 
     cex = 0.5,
     nodePar = 1)
```
