---
title: "Appendix 6b) Model diagnostics (.R)"
author: "Vignette Author"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warnings = F
)
```

For your independent project you need to carry out relevant model diagnostics.  If your major analysis only requires t-tests, you should re-fit at least one of your t-tests a linear model using lm() and check for normality and homogeneity of variance using relevant techniques.

The following script is just a place holder.  For full information on model diangostics see the the (diagnostics tutorial](https://brouwern.github.io/mammalsmilk/articles/g-linear_reg_diagnostics.html) in the [mammalsmilk](https://brouwern.github.io/mammalsmilk/index.html) package.


In this the mammalsmilkRA repository I have split up the analysis into several subscripts:

* analysis
* diagnostics
* figure

These can be combined if you want.


## Preliminaries

```{r}
library(dplyr)  # for exploratory analyses
library(ggpubr) # plotting using ggplto2
library(cowplot)
library(lme4)
library(arm)
library(stringr) 
library(bbmle)
library(plotrix) ##std.error function for SE
library(psych)
library(here)
```




### Load data in R
```{r}
file. <- "Appendix-2-Analysis-Data_mammalsmilkRA.csv"
path. <- here("/inst/extdata/",file.)

milk <- read.csv(path., skip = 3)


genus_spp_subspp_mat <- milk$spp %>% str_split_fixed(" ", n = 3)

milk$genus <- genus_spp_subspp_mat[,1]
```


### Check input

```{r}
head(milk)
tail(milk)
summary(milk)
```

