---
title: "Results - model equation"
author: "Vignette Author"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```



## Introduction

Zuur and Ieno (2016) encourage authors to present the model equation associated with their analyses ("Step 5: Present the statistical model").  This is a very important aspect of an analysis but requires you to know how to interpret R's output and to understand what the underlying relationships which your "Statistical test" implies.

Zuur & Ieno (2016) provide some examples, but they are for fairly complex models.  I am working on simpler examples.

Here are some thoughts on understanding and implementing this task


## Simple regression: use ggpubr!

For simple regression situations we can get ggpubr to come up with the equation for us.

Some packages and data
```{r}
library(ggpubr)
library(MASS)
data(mammals)
```

A simple regression model
```{r}
mod <- lm(brain ~ body, data = mammals)
```

Look at the output
```{r}
summary(mod)
```

The intercept is 91 and the slope is 0.97

This is an equation for

brain ~ body
brain ~ intercept + slope*body
brain ~ 91 + 0.97*body
```{r}
ggscatter(data = mammals,
       y = "brain",
       x = "body", 
       add = "reg.line") +
  stat_regline_equation()
```



## References on regession and regression equations

For general information on how to interpret regression equations see these papers

Boldina & Beninger.  2016.  Strengthening statistical usage in marine ecology: Linear regression.  [Journal of Experimental Marine Biology & Ecology](https://www.sciencedirect.com/science/article/pii/S002209811530023X)

Marril.  2004.  Advanced Statistics: Linear Regression, Part I: Simple Linear Regression.  [Emergency Medicine](https://onlinelibrary.wiley.com/doi/abs/10.1197/j.aem.2003.09.005)


Marril 2008.  Advanced Statistics: Linear Regression, Part II: Multiple Linear Regression.  [Emergency Medicine](https://onlinelibrary.wiley.com/doi/abs/10.1197/j.aem.2003.09.006)

Schneider et al 2010.  Linear regression analysis: part 14 of a series on evaluation of scientific publications.  [Deutsches Ärzteblatt.](https://www.aerzteblatt.de/int/archive/article?id=79009)




Zuur & Ieno.  2016.  A protocol for conducting and presenting results of regression‐type analyses
https://besjournals.onlinelibrary.wiley.com/doi/full/10.1111/2041-210X.12577

