---
title: "z) Creating this package"
author: "Nathan Brouwer"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

## Introduction

The following code outlines the key steps I did, principally using the usethis package.


## Packages

I use these packages for setting things up
```{r, eval = F}
library(devtools)
library(usethis)
library(here)
```

## Create repository

I create a shroom git repo then cloned into it by hand by creating an RStudio Project.

## Set up package with usethis

I then used create_package() within my project directory to build the basic package infrastructure  This overights the initial project (after a handy prompt in the console).
```{r eval = F}
create_package(path = getwd())
```

### Create vignette infrastructure

Create a vignette infrastructure and dummy vignette
```{r, eval = F}
usethis::use_vignette("Results-model-equation")
```

### Create "readme" and "news" files

```{r, eval = F}
use_readme_rmd()
use_news_md()
```



### Create git ignore

Am not sure how to use this properly yet

```{r, eval = F}
use_git_ignore(".pdf", directory = ".")
```


### Add required packages

```{r, eval = F}
use_package("dplyr", "Imports")
use_package("tidyr", "Imports")
use_package("ggplot2", "Imports")
use_package("nlme", "Imports")
use_package("lme4", "Imports")
use_package("bbmle", "Imports") #bbmle
use_package("RCurl", "Imports")
use_package("GGally", "Imports")
use_package("broom", "Imports")
use_package("arm", "Imports")
use_package("psych", "Imports")
devtools::use_package("stringi", "Imports")
devtools::use_package("stringr", "Imports")
devtools::use_package("ggpubr", "Imports")
devtools::use_package("plotrix", "Imports")
devtools::use_package("here", "Imports")

devtools::use_package("cowplot", "Imports")



devtools::use_package("data.tree", "Imports")
devtools::use_package("treemap", "Imports")
#use_package("igraph", "Imports")

devtools::use_package("PRISMAstatement",
                      "Imports")


devtools::use_package("sjPlot", "Imports")
devtools::use_package("sjlabelled", "Imports")
devtools::use_package("sjmisc", "Imports")

devtools::use_package("visreg", "Imports")


```



### Don't save/load user workspace between sessions

```{r, eval = F}
use_blank_slate()
```


### Use roxygen for documentation
```{r, eval = F}
use_roxygen_md()
```


### Package-level documents
"Adds a dummy .R file that will prompt roxygen to generate basic package-level documentation. "
```{r, eval = F}
use_package_doc()
```


### Use pkgdown

For making front end website

```{r, eval = F}
use_pkgdown()
```



## Set up data

```{r}
dir.create(here("/inst"))
dir.create(here("/inst/extdata"))
```


### Look at data in my extdata file

External data is stored in "/inst/extdata"
```{r, eval = F}
list.files(here::here("/inst/extdata"))
```

#### Raw data

```{r eval = F}
milk_raw <- read.csv(here::here("/inst/extdata",
                    "skibiel_mammalsmilk_raw.csv"))

use_data(milk_raw)
use_r("milk_raw")
```




#### Other data data

Just the fat column
```{r eval = F}
milk <- read.csv(here::here("/inst/extdata",
                    "skibiel_mammalsmilk.csv"))

use_data(milk)
use_r("milk")
```


Just the primates and relatives rows
```{r  eval = F}
milk_primates <- read.csv(here::here("/inst/extdata",
                    "skibiel_primate_fat.csv"))

use_data(milk_primates)
use_r("milk_primates")
```




Just the fat column
```{r  eval = F}
milk_fat <- read.csv(here::here("/inst/extdata",
                    "skibiel_fat.csv"))

use_data(milk_fat)
use_r("milk_fat")
```


## License

http://kbroman.org/pkg_primer/pages/licenses.html

```{r, eval = F}
usethis::use_mit_license(name = "Nathan Brouwer")
```






```{r}
#use_data_raw()
```

### Spell check

use_spell_check


### Other potentially useful usethis functions

use_r() #Create or edit a .R file


use_build_ignore()
Add files to .Rbuildignore




use_version() use_dev_version()
Increment package version



edit_r_profile() 
edit_r_environ() 
edit_r_makevars() 
edit_rstudio_snippets() 
edit_git_config() 
edit_git_ignore()



use_directory()
Use a directory




## devtools functions

"Builds package vignettes using the same algorithm that R CMD build does"

```{r, eval = F}
devtools::build_vignettes()
```

"This uses a fairly rudimentary algorithm where any files in ‘inst/doc’ with a name that exists in ‘vignettes’ are removed."

```{r, eval = F}

clean_vignettes() 
```


## pckgdown functions

```{r, eval = F}
pkgdown::build_site()
```

check_failures
```{r}

```


## Rcmd.exe
"Clean and rebuild"
==> Rcmd.exe INSTALL --preclean --no-multiarch --with-keep.source mammalsmilk

"Install restart"
==> Rcmd.exe INSTALL --no-multiarch --with-keep.source mammalsmilk

## Devtools function


```{r, eval = F}
## "Document"
devtools::document(roclets=c('rd', 'collate', 'namespace'))

## "Load all"
devtools::load_all(".")

## devtools::build_vignettes()

## "Check"
devtools::check()
```




